Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-18 ind_files/05_unification_discussion.tex
Minor phrasing changes

In chapter \ref{chap_infD_processes}, I formulated a `field equation' for the evolution of one-dimensional quantitative traits in populations. To recap, given a one-dimensional quantitative trait that takes values in a trait space $\mathcal{T} \subseteq \mathbb{R}$, I defined the set $\mathcal{M}(\mathcal{T}) =  \left\{\sum_{i=1}^{n}\delta_{x_i} \ | \ n \in \mathbb{N}, x_i \in \mathcal{T}\right\}$, where each $\delta_{x_i}$ is a Dirac mass centered at $x_i$. I then formulated a model in which the population at time $t$ is characterized as a whole by a randomly varying density distribution (`stochastic field') $\nu^{(t)}: \mathcal{M}(\mathcal{T}) \to [0,\infty)$ such that for any subset $A \subseteq \mathcal{T}$, the number of individuals that have trait value in $A$ is given by integrating $\nu^{(t)}$ over $A$. The change of this field is determined entirely by two functionals, $b(x|\nu)$ and $d(x|\nu)$ from $\mathcal{T}$ to $[0,\infty)$, that respectively describe the birth rate and death rate of individuals of type $x$ in a population $\nu$. Under the assumption that there exists a suitable system size parameter $K > 0$, I moved from the space of `number' distribution functions $\mathcal{M}(
\mathcal{T})$ to the space of `density' distribution functions $\mathcal{M}_{K}(\mathcal{T}) =  \left\{\sum_{i=1}^{n}\delta_{x_i}/K \ | \ n \in \mathbb{N}, x_i \in \mathcal{T}\right\}$ via a functional analog of the system size expansion. By appropriately rescaling the birth and death rate functionals, I then determined that $P(\phi,t)$, the probability that the population is described by the distribution $\phi \in \mathcal{M}_{K}(\mathcal{T})$ at time $t$, (approximately) satisfies the `stochastic field equation':
\begin{equation}
\label{disc_functional_field_eqns}
\resizebox{\textwidth}{!}{$\displaystyle
\frac{\partial P}{\partial t}(\phi,t) = \int\limits_{\mathcal{T}}\left[-
\frac{\delta}{\delta\phi(x)} \big{\{} \big{[}\phi(x) w(x|\phi) + \mu Q(x|\phi)\big{]} P(\phi,t) \big{\}} + \frac{1}{2K}\frac{\delta^2}{\delta\phi(x)^2}\big{\{}\big{[}\phi(x) \tau(x|\phi) + \mu Q(x|\phi)\big{]}P(\phi,t)\big{\}}\right]dx$}
\end{equation}
where $w(x|\phi)$, $\tau(x|\phi)$, and $Q(x|\phi)$ are functionals that respectively describe the Malthusian fitness, per-capita turnover rate, and birth rate due to mutations (with mutation rate $\mu$) of type $x \in \mathcal{T}$ individuals (now a continuous variable) in a population $\phi$. We then saw that this equation yields some well-known frameworks of quantitative genetics in the infinite population ($K \to \infty$) limit, thus illustrating consistency with known theories.

>>>>

In chapter \ref{chap_infD_processes}, I formulated a `field equation' for the evolution of one-dimensional quantitative traits in populations. To recap, given a one-dimensional quantitative trait that takes values in a trait space $\mathcal{T} \subseteq \mathbb{R}$, I defined the set $\mathcal{M}(\mathcal{T}) =  \left\{\sum_{i=1}^{n}\delta_{x_i} \ | \ n \in \mathbb{N}, x_i \in \mathcal{T}\right\}$, where each $\delta_{x_i}$ is a Dirac mass centered at $x_i$. I then formulated a model in which the population at time $t$ is characterized as a whole by a randomly varying density distribution (`stochastic field') $\nu^{(t)}: \mathcal{M}(\mathcal{T}) \to [0,\infty)$ such that for any subset $A \subseteq \mathcal{T}$, the number of individuals that have trait value in $A$ is given by integrating $\nu^{(t)}$ over $A$. The change of this field is determined entirely by two functionals, $b(x|\nu)$ and $d(x|\nu)$ from $\mathcal{T} \times \mathcal{M}(\mathcal{T})$ to $[0,\infty)$, that respectively describe the birth rate and death rate of individuals of type $x$ in a population $\nu$. Under the assumption that there exists a suitable system size parameter $K > 0$, I moved from the space of `number' distribution functions $\mathcal{M}(
\mathcal{T})$ to the space of `density' distribution functions $\mathcal{M}_{K}(\mathcal{T}) =  \left\{\sum_{i=1}^{n}\delta_{x_i}/K \ | \ n \in \mathbb{N}, x_i \in \mathcal{T}\right\}$ via a functional analog of the system size expansion. By appropriately rescaling the birth and death rate functionals, I then determined that $P(\phi,t)$, the probability that the population is described by the distribution $\phi \in \mathcal{M}_{K}(\mathcal{T})$ at time $t$, (approximately) satisfies the `stochastic field equation':
\begin{equation}
\label{disc_functional_field_eqns}
\resizebox{\textwidth}{!}{$\displaystyle
\frac{\partial P}{\partial t}(\phi,t) = \int\limits_{\mathcal{T}}\left[-
\frac{\delta}{\delta\phi(x)} \big{\{} \big{[}\phi(x) w(x|\phi) + \mu Q(x|\phi)\big{]} P(\phi,t) \big{\}} + \frac{1}{2K}\frac{\delta^2}{\delta\phi(x)^2}\big{\{}\big{[}\phi(x) \tau(x|\phi) + \mu Q(x|\phi)\big{]}P(\phi,t)\big{\}}\right]dx$}
\end{equation}
where $w(x|\phi)$, $\tau(x|\phi)$, and $Q(x|\phi)$ are functionals that respectively describe the Malthusian fitness, per-capita turnover rate, and birth rate due to mutations (with mutation rate $\mu$) of type $x \in \mathcal{T}$ individuals (now a continuous variable) in a population $\phi$. We then saw that this equation yields some well-known frameworks of quantitative genetics in the infinite population ($K \to \infty$) limit, thus illustrating consistency with known theories.
